---
title: "Big Year Using eBird Data"
author: "Julie Mallon"
date: "5/20/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

![]() big year movie photo

The objective of this code is to create the ultimate big year list where every bird species that lives in the USA is accounted for between Jan 1 - Dec 31. 
We will do this by:
+ generating a reference list of species in the USA that are not vagrants or subspecies
+ accessing eBird's 2019 data using APIs
+ selecting one checklist per day that will get us the closest to our target species count
+ writing an algorithm that prioritizes highly endemic and other rare/cryptic species 

The algorithm will then sequentially fill in missing species inversely related to their frequency. Let's see how quickly we can achieve *the* Big Year!


First we will do this with a single US state - Delaware.


```{r, eval = F, echo = T}
install.packages('rebird')

library(rebird)
library(readr)    
library(dplyr)  

source('myfilepath/ebird_targets_functions.R') 
remotes::install_github("CornellLabofOrnithology/auk")

api_key = 'myapikey'
```


Get current taxonomic list for North and Middle America
```{r}
library (RCurl)

AOS_taxonomy_csv <- getURL("http://checklist.americanornithology.org/taxa.csv?type=charset%3Dutf-8%3Bsubspecies%3Dno%3B")
AOS_taxa <- read.csv (text = AOS_taxonomy_csv)
```

```{r, eval = T}
wd <- 'C:/Users/Julie/GitHub/BigYear'
#library(readr)    
#library(dplyr)  

library(rebird)

source(file.path(wd,'BigYear_functions.R')) 
#remotes::install_github("CornellLabofOrnithology/auk")

api_key = 'vir66vo33jre'
```

Get the eBird taxonomy list
```{r}
ebird_taxa <- ebirdtaxonomy(key = api_key) #default language is English

```

Subset this to birds that are known to reside at least part of the year in the USA
```{r}

```

Get frequencies of birds in Delaware
```{r}
DE_freq <- ebirdfreq("states", 
                         "US-DE", 
                         startyear = 2019, 
                         endyear = 2019, 
                         startmonth = 1, 
                         endmonth = 12, 
                         long = TRUE)
head(DE_freq)
```

Notice that there are many species with 0's for frequencies. Let's remove those.
```{r}
DE_observed <- DE_freq[which(DE_freq$frequency>0),]

DE_ebird_species <- unique(DE_observed$comName)
```

Now we have a working list of all the species that were seen in Delaware in 2019. Note that eBird taxonomy includes groups of birds (if unidentifiable to species), pairs of birds (that are difficult to ID to species) and hybirds! We want to remove those.
```{r}
tail(DE_ebird_species)

nonspecies_remove <- c(grep("sp.", DE_ebird_species),
                       grep("/", DE_ebird_species),
                       grep("(hybrid)", DE_ebird_species))

DE_species_2019 <- DE_ebird_species[-nonspecies_remove]
length(DE_species_2019)
```
`r length(DE_species_2019)` is a lot of species for one of the smallest states in the US!

Filter the frequency data to include only the species in our DE_species_2019 vector
```{r}
DE_observed <- DE_observed[which(DE_observed$comName %in% DE_species_2019),]

```

```{r, eval = F}
# on rebird
ebirdhistorical(loc = 'US-VA-003', date='2019-02-14',max=10)

# non rebird functions
https://api.ebird.org/v2/product/stats/{{regionCode}}/{{y}}/{{m}}/{{d}}
#Get a summary of the number of checklist submitted, species seen and contributors on a given date for a country or region.

#regionCode	The country, subnational1, subnational2 or location code.
#year	The year, from 1800 to the present.
#month	The month, from 1 -12.
#day	The day in the month.

https://api.ebird.org/v2/product/lists/{{regionCode}}/{{y}}/{{m}}/{{d}}
#Get information on the checklists submitted on a given date for a country or region.

#regionCode	The country, subnational1, subnational2 or location code.
#year	The year, from 1800 to the present.
#month	The month, from 1 -12.
#day	The day in the month.
```

```{r, eval = F}

# Base URL for eBird API 2.0
url_base_obs = 'https://ebird.org/ws2.0/data/obs/'

# Create a list to hold the individual dictionaries of observations
observations = NULL

# Loop over the locations of interest and dates of interest
for(loc_id in loc_ids){
  for(d in rng):
  time.sleep(0.5) # time delay
}
ymd = '{}/{}/{}'.format(d.year, d.month, d.day)

# Build the URL
url_obs = url_base_obs + loc_id + '/historic/' + ymd + \
'?rank=mrec&detail=full&cat=species&key=' + api_key
print(url_obs)

# Get the observations for one location and date
obs = requests.get(url_obs)

# Append the new observations to the master list
observations.extend(obs.json())

# Convert the list of dictionaries to a pandas dataframe        
obs_df = pd.DataFrame(observations)

# Check out the structure of the dataframe
print(obs_df.info())

# Check out the first few rows
obs_df.head()

# Export the dataframe to a csv file
obs_df.to_csv("observations.csv", index=False)

```